---
title: "Algal turf and macroalgae productivity from the AIMS LTMP"
author: "Renato"
date: '2022-06-30'
output: pdf_document
---

```{r, message=FALSE}
library(tidyverse)
library(brms)
library(tidybayes)
library(patchwork)
library(PNWColors)
```

## Loading the two meta-analysis models
One developed to predict algal turf productivity from depth,
and the other simply as an average macroalgal productivity (no predictors)

```{r}
macr_prodmod <- readRDS('mods/macro_prod_brms.RDS')
turf_prodmod <- readRDS('mods/turf_prod_brms.RDS')
```

## And loading the AIMS LTMP data for which they will be predicted

```{r}
ltmp_cover <- read.csv('data/AIMS_LTMP_cover.csv') %>%
  mutate(cover=total_cover/100) %>%
    select(-X,-total_cover)

ltmp_depth <- read.csv('data/AIMS_LTMP_depths.csv') %>%
  mutate(REEF_NAME=tolower(gsub(' ','_', REEF_NAME))) %>%
  group_by(REEF_NAME, SITE_NO, REPORT_YEAR, TRANSECT_NO) %>%
  summarise(depth=mean(START_DEPTH), .groups='drop_last')

names(ltmp_depth) <- tolower(names(ltmp_depth))


ltmp <- left_join(
  ltmp_cover,
  ltmp_depth,
  by=c('reef_name','site_no','transect_no','report_year')
) %>%
  pivot_wider(names_from='group_code',
              values_from='cover') %>%
  rename(algal_turf = ta, macroalgae = ma)

```


## Turf prediction coming

```{r}

pred_val <- posterior_epred(turf_prodmod, 
      newdata=ltmp %>% mutate(se_prod=0.1), ndraws=1000)

tot_turf_prod <- ((ltmp$algal_turf * t(pred_val)) * 10000)/1000
## in kg C ha day-1


# aggregating

ltmp_pred <- cbind(ltmp,
      turf_prod_kghaday_med=apply(tot_turf_prod,1,median),
      turf_prod_kghaday_lhd=apply(tot_turf_prod,1,function(x) median_hdci(x)$ymin),
      turf_prod_kghaday_uhd=apply(tot_turf_prod,1,function(x) median_hdci(x)$ymax))

```

## But also macroalgae

```{r}

pred_val2 <- posterior_epred(macr_prodmod, 
      newdata=ltmp %>% mutate(se_prod=0.1), ndraws=1000)

tot_macr_prod <- ((ltmp$macroalgae * t(pred_val2)) * 10000)/1000
## in kg C ha day-1


# aggregating

ltmp_pred <- cbind(ltmp_pred,
      macr_prod_kghaday_med=apply(tot_macr_prod,1,median),
      macr_prod_kghaday_lhd=apply(tot_macr_prod,1,function(x) median_hdci(x)$ymin),
      macr_prod_kghaday_uhd=apply(tot_macr_prod,1,function(x) median_hdci(x)$ymax))

```


### First the time series of turf productivity over time in Moorea
```{r, fig.height=5, fig.width=3.125, warning=FALSE}

ggplot(data=ltmp_pred) +
  geom_line(aes(x=report_year,y=turf_prod_kghaday_med,group=interaction(transect_no,site_no))) +
  geom_smooth(aes(x=report_year,y=turf_prod_kghaday_med), 
              method = 'loess', formula = y ~ x,
              colour='orange', alpha=0.8) +
  #scale_x_continuous(breaks=c(2005,2009,2013,2017)) +
  facet_wrap(~reef_name,nrow=3) +
  ylab(expression(Algal~turf~productivity~'('*kg~C~ha^-1*day^-1*')')) +
  theme_minimal()

```


## How to they relate to each other?

```{r, warning=FALSE}
ggplot(ltmp_pred) +
  geom_linerange(
    aes(xmin=turf_prod_kghaday_lhd,
        xmax=turf_prod_kghaday_uhd,
        y=macr_prod_kghaday_med),
    size=0.1) +
  geom_linerange(
    aes(ymin=macr_prod_kghaday_lhd,
        ymax=macr_prod_kghaday_uhd,
        x=turf_prod_kghaday_med),
    size=0.1) +
  geom_point(
    aes(x=turf_prod_kghaday_med,
        y=macr_prod_kghaday_med,
        colour=report_year)) +
  scale_colour_gradientn(colours=pnw_palette("Bay",10)) +
  scale_y_log10() +
  labs(x=expression(Algal~turf~productivity~'('*kg~C~ha^-1*day^-1*')'),
       y=expression(Macroalgae~productivity~'('*kg~C~ha^-1*day^-1*')')) +
  theme_minimal()

```

## Saving

```{r}
saveRDS(ltmp_pred,'preds/LTMP_TurfMacroalgaeProd_preds.RDS')
```